---
output:
  html_document: default
  pdf_document: default
---
Explore and Summarize Data Project by David Smyth
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
options(repos="https://cran.rstudio.com" )
```

```{r echo=FALSE, Load_the_Data1}
# Load the Data
#getwd()
#list.files()
pLD = read.csv('prosperLoanData.csv')
```

```{r echo=FALSE, Load_the_Data2}
# Explore dataset
#names(pLD)
```

```{r echo=FALSE, Load_the_Data3}
# Explore dataset
summary(pLD)
```

The Prosper Loan Data dataset contains 113,937 loans with 81 variables on each
loan, including loan amount, borrower rate (or interest rate), current loan
status, borrower income, borrower employment status, borrower credit history,
and the latest payment information.

# Univariate Plots Section

```{r echo=FALSE, Univariate_Plots1}
  summary(pLD$CreditGrade)
```

```{r echo=FALSE, warning=FALSE, Univariate_Plots2}
ggplot(aes(x = CreditGrade), data = pLD) +
  stat_count(width = 0.5) +
  scale_y_log10()
```

The Credit rating that was assigned at the time the listing went live.
Applicable for listings pre-2009 period and will only be populated for those
listings. The y-axis is presented on a logarithmic scale.

```{r echo=FALSE, Univariate_Plots3}
  summary(pLD$LoanStatus)
```

```{r echo=FALSE, warning=FALSE, Univariate_Plots4}
ggplot(aes(x = LoanStatus), data = pLD) +
  stat_count(width = 0.5) +
  scale_x_discrete() +
  theme(axis.text.x = element_text(angle=60, hjust=1)) +
  scale_y_log10()
```

The current status of the loan: Cancelled, Chargedoff, Completed, Current,
Defaulted, FinalPaymentInProgress, PastDue. The PastDue status will be
accompanied by a delinquency bucket. The y-axis is presented on a logarithmic
scale.

```{r echo=FALSE, Univariate_Plots5}
  summary(pLD$BorrowerAPR)
```

```{r echo=FALSE, warning=FALSE, Univariate_Plots6}
library(gridExtra)
library(grid)
library(lattice)
grid.arrange( ggplot(aes(x=BorrowerAPR), 
        data = subset(pLD,BorrowerAPR!='NA')) +
  geom_histogram( bins = 30) ,
   ggplot(aes(x=1, y=BorrowerAPR), 
        data = subset(pLD,BorrowerAPR!='NA')) +
  geom_boxplot( )  , nrow =1)
```

The Borrower's Annual Percentage Rate (APR) for the loan. 25 NA's have been
removed in presentation of the data.

```{r echo=FALSE, Univariate_Plots7}
  summary(pLD$BorrowerRate)
```

```{r echo=FALSE, warning=FALSE, Univariate_Plots8}
grid.arrange( ggplot(aes(x=BorrowerRate), 
        data = pLD) +
  geom_histogram( bins = 30) ,
   ggplot(aes(x=1, y=BorrowerRate), 
        data = pLD) +
  geom_boxplot( )  , nrow =1)
```

The Borrower's interest rate for this loan.

```{r echo=FALSE, Univariate_Plots9}
  summary(pLD$ProsperRating..numeric.)
```

```{r echo=FALSE, warning=FALSE, Univariate_Plots10}
ggplot(aes(x = ProsperRating..numeric.),
       data = subset(pLD,ProsperRating..numeric.!='NA')) +
  stat_count(width = 0.5) +
  scale_x_continuous()
```

The Prosper Rating assigned at the time the listing was created: 0 - N/A,
1 - HR, 2 - E, 3 - D, 4 - C, 5 - B, 6 - A, 7 - AA. Applicable for loans
originated after July 2009. 29084 NA's have been removed in presentation of the
data.

```{r echo=FALSE, Univariate_Plots11}
  summary(pLD$ProsperRating..Alpha.)
```

```{r echo=FALSE, warning=FALSE, Univariate_Plots12}
ggplot(aes(x = ProsperRating..Alpha.), data = pLD) +
  stat_count(width = 0.5) +
  scale_x_discrete() +
  theme(axis.text.x = element_text(angle=60, hjust=1))
```

The Prosper Rating assigned at the time the listing was created between AA - HR.
Applicable for loans originated after July 2009.

```{r echo=FALSE, Univariate_Plots13}
  summary(pLD$ProsperScore)
```

```{r echo=FALSE, warning=FALSE, Univariate_Plots14}
ggplot(aes(x = ProsperScore), data = subset(pLD,ProsperScore!='NA'),) +
  stat_count(width = 0.5) +
  scale_x_continuous()
```

A custom risk score built using historical Prosper data. The score ranges from
1-10, with 10 being the best, or lowest risk score. Applicable for loans
originated after July 2009. 29084 NA's have been removed in presentation of the
data.

```{r echo=FALSE, Univariate_Plots15}
  summary(pLD$Occupation)
```

```{r echo=FALSE, warning=FALSE, Univariate_Plots16}
ggplot(aes(x = Occupation), data = pLD) +
  stat_count(width = 0.5) +
  scale_x_discrete() +
  theme(axis.text.x = element_text(angle=60, hjust=1)) +
  scale_y_log10()
```

The Occupation selected by the Borrower at the time they created the listing.
The y-axis is presented on a logarithmic scale.

```{r echo=FALSE, Univariate_Plots17}
  summary(pLD$EmploymentStatusDuration)
```

```{r echo=FALSE, warning=FALSE, Univariate_Plots18}
ggplot(aes(x = EmploymentStatusDuration),
       data = subset(pLD,EmploymentStatusDuration!='NA')) +
  geom_histogram(binwidth = 10) +
    coord_cartesian(xlim = c(0, 600))
```

The length in months of the employment status at the time the listing was
created. The x-axis's upper range has been limited to 600 removing max outlier
at 755.00. 7625 NA's have been removed in presentation of the data.

```{r echo=FALSE, Univariate_Plots19}
  summary(pLD$IsBorrowerHomeowner)
```

```{r echo=FALSE, warning=FALSE, Univariate_Plots20}
ggplot(aes(x = IsBorrowerHomeowner), data = pLD) +
  geom_bar() +
  scale_x_discrete()
```

A Borrower will be classified as a homeowner if they have a mortgage on their
credit profile or provide documentation confirming they are a homeowner.

```{r echo=FALSE, Univariate_Plots21}
  summary(pLD$CreditScoreRangeLower)
```

```{r echo=FALSE, warning=FALSE, Univariate_Plots22}
ggplot(aes(x = CreditScoreRangeLower), data = subset(pLD,CreditScoreRangeLower!='NA')) +
  geom_histogram(binwidth = 10) +
  coord_cartesian(xlim = c(450, 900))
```

The lower value representing the range of the borrower's credit score as
provided by a consumer credit rating agency. The x-axis's upper and lower range
has been limited to 450 to 900 removing min outlier at 0.0. 591 NA's have been
removed in presentation of the data.

```{r echo=FALSE, Univariate_Plots23}
  summary(pLD$CreditScoreRangeUpper)
```

```{r echo=FALSE, warning=FALSE, Univariate_Plots24}
ggplot(aes(x = CreditScoreRangeUpper),
       data = subset(pLD,CreditScoreRangeUpper!='NA')) +
  geom_histogram(binwidth = 10) +
  coord_cartesian(xlim = c(450, 900))
```

The upper value representing the range of the borrower's credit score as
provided by a consumer credit rating agency. The x-axis's upper and lower range
has been limited to 450 to 900 removing min outlier at 19.0. 591 NA's have been
removed in presentation of the data.

```{r echo=FALSE, Univariate_Plots25}
  summary(pLD$CurrentDelinquencies)
```

```{r echo=FALSE, warning=FALSE, Univariate_Plots26}
ggplot(aes(x = CurrentDelinquencies),
       data = subset(pLD,CurrentDelinquencies!='NA')) +
  geom_histogram(binwidth = 1) +
    coord_cartesian(xlim = c(0, 20))
```

Number of accounts delinquent at the time the credit profile was pulled. The
x-axis's upper range has been limited to 20 removing max outlier at 83.0000. 697
NA's have been removed in presentation of the data.

```{r echo=FALSE, Univariate_Plots27}
  summary(pLD$AmountDelinquent)
```

```{r echo=FALSE, warning=FALSE, Univariate_Plots28}
ggplot(aes(x = AmountDelinquent), data = subset(pLD,AmountDelinquent!='NA')) +
  geom_histogram(binwidth = 10) +
    scale_x_sqrt() +
    coord_cartesian(xlim = c(0, 50000))
```

Dollars delinquent at the time the credit profile was pulled. The x-axis's upper
range has been limited to 50000 removing max outlier at 463881.0. 7622 NA's have
been removed in presentation of the data.

```{r echo=FALSE, Univariate_Plots29}
  summary(pLD$BankcardUtilization)
```

```{r echo=FALSE, warning=FALSE, Univariate_Plots30}
ggplot(aes(x = BankcardUtilization),
       data = subset(pLD,BankcardUtilization!='NA')) +
  geom_histogram(binwidth = .1) +
  coord_cartesian(xlim = c(0, 2.0))
```

The percentage of available revolving credit that is utilized at the time the
credit profile was pulled. The x-axis's upper range has been limited to 2.0
removing max outlier at 5.950. 7604 NA's have been removed in presentation of
the data.

```{r echo=FALSE, Univariate_Plots31}
  summary(pLD$DebtToIncomeRatio)
```

```{r echo=FALSE, warning=FALSE, Univariate_Plots32}
ggplot(aes(x = DebtToIncomeRatio), data = subset(pLD,DebtToIncomeRatio!='NA')) +
  geom_histogram(binwidth = .1) +
    coord_cartesian(xlim = c(0, 1.5))
```

The debt to income ratio of the borrower at the time the credit profile was
pulled. This value is Null if the debt to income ratio is not available. This
value is capped at 10.01 (any debt to income ratio larger than 1000% will be
returned as 1001%). The x-axis's upper range has been limited to 1.5
removing max outlier at 10.010. 8554 NA's have been
removed in presentation of the data.

```{r echo=FALSE, Univariate_Plots33}
  summary(pLD$IncomeRange)
```

```{r echo=FALSE, warning=FALSE, Univariate_Plots34}
ggplot(aes(x = IncomeRange), data = pLD) +
  geom_bar() +
  theme(axis.text.x = element_text(angle=60, hjust=1))
```

The income range of the borrower at the time the listing was created.

```{r echo=FALSE, Univariate_Plots35}
  summary(pLD$IncomeVerifiable)
```

```{r echo=FALSE, warning=FALSE, Univariate_Plots36}
ggplot(aes(x = IncomeVerifiable), data = pLD) +
  geom_bar()
```

The borrower indicated they have the required documentation to support their
income.

```{r echo=FALSE, Univariate_Plots37}
  summary(pLD$StatedMonthlyIncome)
```
```{r echo=FALSE, warning=FALSE, Univariate_Plots38}
ggplot(aes(x = StatedMonthlyIncome), data = pLD) +
  geom_histogram(binwidth = 10) +
  scale_x_sqrt() +
  coord_cartesian(xlim = c(0, 50000))
```

The monthly income the borrower stated at the time the listing was created. The
x-axis is presented on a square root scale and has been limited to 50000
removing max outlier at 1750003.

```{r echo=FALSE, Univariate_Plots39}
  summary(pLD$LoanOriginalAmount)
```

```{r echo=FALSE, warning=FALSE, Univariate_Plots40}
ggplot(aes(x = LoanOriginalAmount), data = pLD) +
  geom_histogram(binwidth = 1) +
  scale_x_sqrt() +
  scale_y_log10()
```

The origination amount of the loan. The y-axis is presented on a logarithmic
scale, the x-axis is presented on a square root scale.

```{r echo=FALSE, Univariate_Plots41}
  summary(pLD$MonthlyLoanPayment)
```

```{r echo=FALSE, warning=FALSE, Univariate_Plots42}
ggplot(aes(x = MonthlyLoanPayment), data = pLD) +
    geom_histogram(binwidth = 10) +
    coord_cartesian(xlim = c(0, 1500))
```

The scheduled monthly loan payment. The x-axis's upper range has been limited to
1500 removing max outlier at 2251.5.

# Univariate Analysis

### What is the structure of your dataset?

This dataset contains 113,937 loans with 81 variables on each loan, including
loan amount, borrower rate (or interest rate), current loan status, borrower
income, borrower employment status, borrower credit history, and the latest
payment information.

### What is/are the main feature(s) of interest in your dataset?

The main feature of interest will be ProsperScore, a score of risk on a scale of
1 to 10; how it is likely determined, and what factors play a hand in
influencing it.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

Features indicating the customer's ability to take on debt, such as Occupation,
CreditScoreRange, DebtToIncomeRatio, etc.

### Did you create any new variables from existing variables in the dataset?

No new variables have been created in the investigation of this dataset, 81 are
more than enough!

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the \
form of the data? If so, why did you do this?

In general all features were investigated utilizing histograms of various bin
widths and sizes, some scaling of the x-axis using scale_x_sqrt() and some
scaling of the y-axis using  scale_y_log10(). x.element_text was used to
facilitate the reading of tags and coord_cartesian() and x_lim was used to limit
the x-axis. Otherwise, geom_bar() was used to indicate True or False as with
IncomeVerifiable or IsBorrowerHomeowner.

# Bivariate Plots Section

```{r echo=FALSE, Bivariate_Plots1}
summary(pLD$EmploymentStatus)
```

```{r echo=FALSE, warning=FALSE, Bivariate_Plots2}
qplot(x = EmploymentStatus, y = ProsperScore,
      data = pLD,
      geom = 'boxplot')
```

Bivariate boxplot plotting ProsperScore on the y-axis as a function of
EmploymentStatus. Curiously, Self-employed garners a lower ProsperScore than Not
employed as evidenced by a comparison of each of the medians in the
interquartile range of the associated boxplots.

```{r echo=FALSE, Bivariate_Plots3}
summary(pLD$IsBorrowerHomeowner)
```

```{r echo=FALSE, warning=FALSE, Bivariate_Plots4}
qplot(x = IsBorrowerHomeowner, y = ProsperScore,
      data = pLD,
      geom = 'boxplot')
```

Bivariate boxplot plotting ProsperScore as a function of IsBorrowerHomeowner,
as evidenced by the nearly identical medians in both the boxplot's interquartile
range, owning a home has no appreciable effect upon one's ProsperScore.

```{r echo=FALSE, Bivariate_Plots5}
summary(pLD$CreditScoreRangeLower)
```

```{r echo=FALSE, warning=FALSE, Bivariate_Plots6}
ggplot(data = subset(pLD,CreditScoreRangeLower!='NA'),
       aes(CreditScoreRangeLower, ProsperScore)) +
  geom_boxplot(aes(group = cut_width(CreditScoreRangeLower, .25)))
```

Bivariate boxplot plotting ProsperScore as a function of
CreditScoreRangeLower, as evidenced by the median in CreditScoreRangeLower
boxplot's interquartile range, CreditScoreRangeLower is positively correlated
with ProsperScore. 591 NA's have been removed in presentation of the data.

```{r echo=FALSE, Bivariate_Plots7}
summary(pLD$DebtToIncomeRatio)
```

```{r echo=FALSE, warning=FALSE, Bivariate_Plots8}
 ggplot(aes(factor(ProsperScore), 
            DebtToIncomeRatio), 
        data = subset(pLD,ProsperScore!='NA')) +
  geom_jitter( alpha = .1)  + 
  geom_boxplot( alpha = .5,color = 'blue')+ 
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)  +
   ylim(0,1)
```
Bivariate multi-plot of DebtToIncomeRatio as a function of ProsperScore with
red point stat_summary indicating the ProsperScore mean. ProsperScore is
negatively correlated with DebtToIncomeRatio.

```{r echo=FALSE, Bivariate_Plots9}
summary(pLD$IncomeRange)
```

```{r echo=FALSE, warning=FALSE, Bivariate_Plots10}
ggplot(pLD, aes(ProsperScore, IncomeRange)) +
  geom_boxplot(aes(group = cut_width(ProsperScore, 0.25)))
```

Bivariate boxplot plotting ProsperScore as a function of IncomeRange, within
$25,000 - 74,999 IncomeRange, ProsperScore appears uniformly distributed.

```{r echo=FALSE, Bivariate_Plots11}
summary(pLD$StatedMonthlyIncome)
```

```{r echo=FALSE, warning=FALSE, Bivariate_Plots12}
ggplot(pLD, aes(ProsperScore, StatedMonthlyIncome)) +
  geom_boxplot(aes(group = cut_width(ProsperScore, 0.25))) +
  scale_y_log10()
```

Bivariate boxplot plotting ProsperScore as a function of StatedMonthlyIncome,
on a logarithmic scale, approaching 1e+04 StatedMonthlyIncome, ProsperScore
appears uniformly distributed.

```{r echo=FALSE, warning=FALSE, Bivariate_Plots13}
 ggplot(aes(factor(ProsperScore), 
            StatedMonthlyIncome), 
        data = subset(pLD,ProsperScore!='NA')) +
  geom_jitter( alpha = .1)  + 
  geom_boxplot( alpha = .5,color = 'blue')+ 
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)  +
  scale_y_log10()
```

Bivariate multi-plot of StatedMonthlyIncome as a function of ProsperScore with
red point stat_summary indicating the ProsperScore mean. ProsperScore is
uniformly distributed approaching 1e+04 on a logarithmic scale.

```{r echo=FALSE, Bivariate_Plots14}
summary(pLD$MonthlyLoanPayment)
```

```{r echo=FALSE, warning=FALSE, Bivariate_Plots15}
ggplot(pLD, aes(ProsperScore, MonthlyLoanPayment)) +
  geom_boxplot(aes(group = cut_width(ProsperScore, 0.25)))
```

Bivariate boxplot plotting MonthlyLoanPayment as a function of ProsperScore,
ProsperScore appears slightly positively correlated below 500.

```{r echo=FALSE, warning=FALSE, Bivariate_Plots16}
 ggplot(aes(factor(ProsperScore), 
            MonthlyLoanPayment), 
        data = subset(pLD,ProsperScore!='NA')) +
  geom_jitter( alpha = .1)  + 
  geom_boxplot( alpha = .5,color = 'blue')+ 
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4)
```

Bivariate multi-plot of MonthlyLoanPayment as a function of ProsperScore with
red point stat_summary indicating the ProsperScore mean. ProsperScore appears
slightly positively correlated below 500.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

Full-time EmploymentStatus seems to garner a greater ProsperScore than Not
employed yet Not employed garners a greater ProsperScore than Self-employed.
IsBorrowerHomeowner seems to have little bearing on ProsperScore which is
uniformly distributed between True and False. CreditScoreRangeLower, especially
above 800, is positively correlated with ProsperScore. A low DebtToIncomeRatio
is initially negatively correlated to ProsperScore but then vacillates it rises
above 2.5. IncomeRange is positively correlated with ProsperScore above 50k. In
general, StatedMonthlyIncome seems uniformly distributed but with outliers of
greater than 1,500,000 having lower scores than others of StatedMonthlyIncome
250k or much less. MonthlyLoanPayment is slightly positively correlated. 

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

The most interesting relationships observed were that Not employed seems to
garner a greater ProsperScore than Self-employed, which seems counter-intuitive,
and also that DebtToIncomeRatio fluctuates in correlation to ProsperScore as it
rises above 2.5.

### What was the strongest relationship you found?

IsBorrowerHomeowner seems definitely to have no bearing whatsoever on
ProsperScore.

# Multivariate Plots Section

```{r echo=FALSE, warning=FALSE, Multivariate_Plots1}
d <- ggplot(data = subset(pLD,occupation='Computer Programmer'), 
            aes(IsBorrowerHomeowner, ProsperScore))
d + geom_bin2d()
```

Multivariate plot of ProsperScore as a function of IsBorrowerHomeowner for
Computer Programmer occupation, as evidenced by nearly identical shading across
almost all color bars, IsBorrowerHomeowner appears to have no correlation to
ProsperScore. 

```{r echo=FALSE, warning=FALSE, Multivariate_Plots2}
d <- ggplot(data = subset(pLD,occupation='Doctor'), 
            aes(IsBorrowerHomeowner, ProsperScore))
d + geom_bin2d()
```

Multivariate plot of ProsperScore as a function of IsBorrowerHomeowner for
Doctor occupation, as evidenced by nearly identical shading across almost
all color bars, IsBorrowerHomeowner appears to have no correlation to
ProsperScore. 

```{r echo=FALSE, warning=FALSE, Multivariate_Plots3}
d <- ggplot(data = subset(pLD,occupation='Tradesman - Plumber'), 
            aes(IsBorrowerHomeowner, ProsperScore))
d + geom_bin2d()
```

Multivariate plot of ProsperScore as a function of IsBorrowerHomeowner for
Tradesman - Plumber occupation, as evidenced by nearly identical shading across
almost all color bars, Tradesman - Plumber appears to have no correlation to
ProsperScore.

```{r echo=FALSE, warning=FALSE, Multivariate_Plots4}
ggplot(aes(y = CreditScoreRangeLower, 
           x = factor(ProsperScore) ), 
       data = subset (pLD, Occupation=='Computer Programmer' | 
                        Occupation=='Doctor' | 
                        Occupation== 'Tradesman - Plumber')) + 
   geom_boxplot( aes(fill= Occupation))  +
  scale_fill_brewer(type='seq')
```

Multivariate plot of CreditScoreRangeLower as a function of ProsperScore for
Computer Programmer, Doctor and Tradesman - Plumber occupations,
CreditScoreRangeLower appears slightly positively correlated with ProsperScore
for all three occupations.

```{r echo=FALSE, warning=FALSE, Multivariate_Plots5}
ggplot(aes(y = CreditScoreRangeUpper, 
           x = factor(ProsperScore) ), 
       data = subset (pLD,Occupation=='Computer Programmer' | 
                        Occupation=='Doctor' | 
                        Occupation== 'Tradesman - Plumber')) + 
   geom_boxplot( aes(fill= Occupation))  +
  scale_fill_brewer(type='seq')
```

Multivariate plot of CreditScoreRangeUpper as a function of ProsperScore for
Computer Programmer, Doctor and Tradesman - Plumber occupations,
CreditScoreRangeUpper appears slightly positively correlated with ProsperScore
for all three occupations.

```{r echo=FALSE, warning=FALSE, Multivariate_Plots6}
ggplot(aes(y = DebtToIncomeRatio, 
           x = factor(ProsperScore) ), 
       data = subset (pLD,Occupation=='Computer Programmer' | 
                        Occupation=='Doctor' | 
                        Occupation== 'Tradesman - Plumber')) + 
   geom_boxplot( aes(fill= Occupation))  +
  scale_fill_brewer(type='seq') +
  scale_y_log10()
```

Multivariate plot of DebtToIncomeRatio as a function of ProsperScore for
Computer Programmer, Doctor and Tradesman - Plumber occupations,
DebtToIncomeRatio appears somewhat negatively correlated with
ProsperScore for all three occupations with many outliers below the given
median.

```{r echo=FALSE, warning=FALSE, Multivariate_Plots7}
ggplot(aes(y = StatedMonthlyIncome, 
           x = factor(ProsperScore) ), 
       data = subset (pLD,Occupation=='Computer Programmer' | 
                        Occupation=='Doctor' | 
                        Occupation== 'Tradesman - Plumber')) + 
   geom_boxplot( aes(fill= Occupation))  +
  scale_fill_brewer(type='seq')
```

Multivariate plot of StatedMonthlyIncome as a function of ProsperScore for
Computer Programmer, Doctor and Tradesman - Plumber occupations,
StatedMonthlyIncome appears slightly positively correlated with ProsperScore for
all three occupations with a few outliers above the given median.

```{r echo=FALSE, warning=FALSE, Multivariate_Plots8}
ggplot(aes(y = MonthlyLoanPayment, 
           x = factor(ProsperScore) ), 
       data = subset (pLD,Occupation=='Computer Programmer' | 
                        Occupation=='Doctor' | 
                        Occupation== 'Tradesman - Plumber')) + 
   geom_boxplot( aes(fill= Occupation))  +
  scale_fill_brewer(type='seq')
```

Multivariate plot of MonthlyLoanPayment as a function of ProsperScore for
Computer Programmer, Doctor and Tradesman - Plumber occupations,
MonthlyLoanPayment appears somewhat positively correlated  with ProsperScore for
all three occupations with some outliers above the given median.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

IsBorrowerHomeowner appears evenly divided between all three occupational
classes.

For all three occupations, ProsperScore seems to roughly correspond to
CreditScore based on the CreditScoreRangeUpper and CreditScoreRangeLower plots.

DebtToIncomeRatio appears somewhat negatively correlated with
ProsperScore for all three occupations.

StatedMonthlyIncome appears slightly positively correlated with ProsperScore for
all three occupations.

MonthlyLoanPayment appears somewhat positively correlated  with ProsperScore for
all three occupations.

### Were there any interesting or surprising interactions between features?

That IsBorrowerHomeowner seems evenly divided between all occupational classes
is somewhat surprising as one would expect a greater percentage of Doctors for
instance to own a home or that Doctor's with a previous home not to need a loan.
Perhaps for a second home?

------

# Final Plots and Summary

### ProsperScore as a function of EmploymentStatus
```{r echo=FALSE, warning=FALSE, Plot_One}
qplot(x = EmploymentStatus, y = ProsperScore,
      data = subset(pLD,ProsperScore!='NA'),
      geom = 'boxplot')
```

Bivariate boxplot plotting ProsperScore on the y-axis as a function of
EmploymentStatus. Curiously, Self-employed garners a lower ProsperScore than Not
employed as evidenced by the lesser median in the interquartile range of the
Self-employed boxplot in comparison with the greater median in the interquartile
range of the Not employed boxplot.

### ProsperScore as a function of IsBorrowerHomeowner
```{r echo=FALSE, warning=FALSE, Plot_Two}
qplot(x = IsBorrowerHomeowner, y = ProsperScore,
      data = pLD,
      geom = 'boxplot')
```

Bivariate boxplot plotting ProsperScore as a function of IsBorrowerHomeowner,
as evidenced by the nearly identical medians in both the boxplot's interquartile
range, owning a home has no appreciable effect upon one's ProsperScore.

### ProsperScore as a function of DebtToIncomeRatio and Occupation

```{r echo=FALSE, warning=FALSE, Plot_Three}
ggplot(aes(y = DebtToIncomeRatio, 
           x = factor(ProsperScore) ), 
       data = subset (pLD,Occupation=='Computer Programmer' | 
                        Occupation=='Doctor' | 
                        Occupation== 'Tradesman - Plumber')) + 
   geom_boxplot( aes(fill= Occupation))  +
  scale_fill_brewer(type='seq') +
  scale_y_log10()
```
 
Multivariate plot of DebtToIncomeRatio as a function of ProsperScore for
Computer Programmer, Doctor and Tradesman - Plumber occupations,
DebtToIncomeRatio appears somewhat negatively correlated with
ProsperScore for all three occupations with many outliers below the given
median.

------

# Reflection

The Prosper Loan Data is large and incorporates many variables which makes it
difficult to decide what to focus on but after much study and reflection,
ProsperScore presented itself as deserving a more through examination... What is
ProsperScore, and if, presumably loans are to be granted based upon it, given
it's proprietary nature, how is it determined.

Exploring the Prosper Loan Data in regards to ProsperScore met with a few
difficulties due to the sprawling number of variables and their size, for
instance, plotting the entire number of Occupations listed, surpassed ggplot's
ability to display and I was forced to settle upon three Occupations to grossly
represent, the blue collar, white collar and professional realms so as to more
simply and clearly delineate possible relationships.

Ultimately, it seems that ProsperScore is a complex calculation that cannot be
easily pinned down via an analysis of three variables as no doubt each of the
Prosper dataset's 81 variables play a part in a much more complex determination
of a customer's final Score. This conclusion is bolstered by the somewhat
counter-intuitive findings garnered by the dataset's exploration and analysis in
R: for instance, that someone Self-employed would receive a lower ProsperScore
than someone Not employed, or that IsBorrowerHomeowner being True receives no
preference in ProsperScore, or as an outlier, a Doctor's ProsperScore can rise
the higher his DebtToIncomeRatio climbs. How these variables are offset by other
heretofore unexamined variables can be a focus of future work within the
dataset.